---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# snic <a id="top"></a>

<!-- badges: start -->
[![R-CMD-check](https://github.com/rolfsimoes/snic/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/rolfsimoes/snic/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Efficient superpixel segmentation for multi-band imagery using the Simple Non-Iterative Clustering (SNIC) algorithm. The package wraps a C++ implementation with an ergonomic R interface, integrates with `terra` for raster workflows, and provides helpers for seed placement, plotting, and reproducibility.

## Installation

``` r
# install.packages("pak")
pak::pak("rolfsimoes/snic")

# alternatively with remotes
# remotes::install_github("rolfsimoes/snic")
```

The `terra` package is suggested for raster support and required for most of the plotting utilities demonstrated below.

## Highlights

- Implements SNIC with a fast C++ core exposed to R
- Works with in-memory arrays or `terra::SpatRaster` objects
- Offers multiple seeding strategies (`snic_rect_grid()`, `snic_diamon_grid()`, `snic_hexagonal_grid()`, `snic_random_grid()`) and interactive placement via `snic_manual_grid()`
- Includes ready-to-plot utilities (`snic_plot()`) for quick inspection of inputs, seeds, and resulting segments
- Ships with a Sentinel-2 subset (`system.file("S2-20LMR", package = "snic")`) for reproducible examples and tests

## Quick start

The example below demonstrates a typical SNIC workflow with the bundled Sentinel-2 subset. The chunk is marked `eval = FALSE` to avoid heavy computation when knitting.

```{r quick-start, eval = TRUE}
load_snic <- function() {
  if (requireNamespace("snic", quietly = TRUE)) {
    exports <- tryCatch(
      getNamespaceExports("snic"),
      error = function(e) character()
    )
    if ("snic_rect_grid" %in% exports) {
      return(library(snic))
    }
  }
  if (requireNamespace("pkgload", quietly = TRUE)) {
    return(pkgload::load_all(export_all = FALSE, helpers = FALSE, attach_testthat = FALSE))
  }
  stop("Install the development version (via pkgload) or the released snic package to run this example.", call. = FALSE)
}

load_snic()

library(terra)

# Sentinel-2 subset packaged with snic
data_dir <- system.file("S2-20LMR", package = "snic", mustWork = TRUE)
bands <- c("B02", "B04", "B08", "B12")
paths <- file.path(
  data_dir,
  sprintf("S2_20LMR_%s_20220630.tif", bands)
)

s2 <- terra::rast(paths)

# Downsample to speed up the example
s2_small <- terra::aggregate(s2, fact = 5)

# Seed generation and segmentation
spacing <- 8L
seeds <- snic_rect_grid(s2_small, spacing = spacing, padding = 0L)
segments <- snic(s2_small, seeds = seeds, compactness = 0.1)

# Visualise RGB composite with superpixel boundaries
snic_plot(
  s2_small,
  r = 4, g = 3, b = 1,
  stretch = "lin",
  seg = segments,
  seg_plot_args = list(border = "#FFFF00", col = NA, lwd = 0.6)
)
```

For quick inspections from the console you can also rely on `snic_plot()`:

```{r snic-plot, eval = TRUE}
snic_plot(
  s2_small,
  r = 4, g = 3, b = 1,
  seeds = seeds,
  seg = segments
)
```

## Seed strategies

All seed generators share the same arguments and return a two-column matrix (`r`, `c`) ready for `snic()`. Use `snic_count_seeds()` to estimate the expected number of superpixels for a given spacing before running the segmentation.

```{r seeding, eval = TRUE}
spacing <- c(20L, 20L)

rect <- snic_rect_grid(s2_small, spacing = spacing)
hex  <- snic_hexagonal_grid(s2_small, spacing = spacing)
rand <- snic_random_grid(s2_small, spacing = spacing)

snic_count_seeds(s2_small, spacing = spacing)
```

When automatic strategies are not enough, `snic_manual_grid()` launches an interactive session (requires an interactive R session with plotting enabled) that lets you refine or extend an existing seed set.

## Scripted demos

The repository provides a ready-to-run script demonstrating a full segmentation workflow for the Rondônia Sentinel-2 stack:

- `inst/scripts/run_snic_rondonia.R`: command-line script that loads the packaged data, runs SNIC, and visualises the result.

Run it with:

``` bash
Rscript inst/scripts/run_snic_rondonia.R
```

## Contributing

Bug reports, feature requests, and pull requests are welcome in the [issue tracker](https://github.com/rolfsimoes/snic/issues). When proposing changes, please run `R CMD check` locally to ensure the package remains stable.
