---
title: "SNIC Segmentation Pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SNIC Segmentation Pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.width = 6,
  fig.height = 5
)
library(snic)
```

## Pipeline overview

The Simple Non-Iterative Clustering (SNIC) algorithm grows superpixels from an initial set of seed pixels. A typical workflow is:

1. Load or prepare an image cube with one or more spectral bands.
2. Place seeds with one of the `snic_grid()` helpers.
3. Call `snic()` to segment the image using the chosen compactness.
4. Inspect the result or convert it to polygons for further analysis.

This vignette walks through that pipeline by working with the Sentinel-2 subset shipped with the package and highlights the most important parameters.

## Loading imagery with terra

```{r load-imagery}
if (requireNamespace("terra", quietly = TRUE)) {
  library(terra)

  data_dir <- system.file("S2-20LMR", package = "snic", mustWork = TRUE)
  band_names <- c("S2_20LMR_B02_20220630.tif",
                  "S2_20LMR_B04_20220630.tif",
                  "S2_20LMR_B08_20220630.tif",
                  "S2_20LMR_B12_20220630.tif")
  s2 <- rast(file.path(data_dir, band_names))

  # Aggregate raster for lightweight examples
  s2_small <- aggregate(s2, fact = 6)
  s2_small
} else {
  stop("Package 'terra' is required for this example.")
}
```

`terra::rast()` reads the reference bands and stores them as a `SpatRaster`.

## Understanding `snic()` parameters

The `snic()` interface is intentionally minimal:

- `img`: array or `SpatRaster` containing the bands to segment.
- `seeds`: pixel coordinates that anchor cluster growth. They must be supplied explicitly, which keeps the results reproducible.
- `compactness`: non-negative value that balances spectral similarity and spatial proximity. Smaller numbers privilege spectral contrast, whereas larger ones grow rounder segments.

The default compactness value of `1` offers a good starting point on reflectance-scaled data. Images with different value ranges can be re-scaled before segmentation (see `?terra::scale`).

## Seed placement with `snic_grid_rect()`

`snic_grid_rect()` builds a uniform rectangular grid of seeds. Two arguments control the layout:

- `spacing`: expected distance (in pixels) between neighboring seeds. A single value applies to rows and columns equally; a length-two vector defines anisotropic spacing.
- `padding`: margin from the image borders where seeds are not placed. The default (`spacing / 2`) centers the grid inside the scene.

```{r rect-grid}
spacing_demo <- 30L
seeds_default <- snic_grid_rect(s2_small, spacing = spacing_demo)
head(seeds_default)
```

As `spacing` increases the number of seeds decreases. Use `snic_count_seeds()` to estimate how many seeds will be created without actually generating them:

```{r count-seeds}
seed_total <- snic_count_seeds(
  s2_small,
  spacing = spacing_demo,
  padding = spacing_demo / 2
)
seed_total
```

Padding creates inner margins and can be set independently along rows and columns:

```{r rect-grid-padding}
seeds_asymmetric <- snic_grid_rect(
  s2_small,
  spacing = c(30, 20),
  padding = c(10, 5)
)
nrow(seeds_asymmetric)
```

## End-to-end segmentation example

```{r run-snic}
comp <- 0.25
seg_rect <- snic(
  s2_small,
  seeds = seeds_default,
  compactness = comp
)
seg_rect
```

Overlay the segmentation to inspect the results:

```{r plot-snic, fig.height=4.5, fig.width=7}
snic_plot(
  s2_small,
  r = 3, g = 2, b = 1,
  stretch = "lin",
  seeds = seeds_default,
  seg = seg_rect,
  seg_plot_args = list(border = "#FFFF00", col = NA, lwd = 0.6),
  seeds_plot_args = list(pch = 3, col = "#00FFFF", cex = 0.7)
)
```

The compactness parameter gently constrains the shapes of the superpixels. Combining a carefully chosen spacing and padding with compactness controls how detailed the segmentation becomes. Dedicated vignettes cover parameter tuning, seed placement strategies, non-Earth-observation imagery, and parallel processing for more advanced workflows.
